---
title: 'Assignment #5'
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
# SEE modeldata package for new datasets
library(tidyverse)         # for graphing and data cleaning
library(tidymodels)        # for modeling
library(ranger)            # for random forest - will need for shiny app
library(lubridate)         # for date manipulation
library(themis)            # for up and downsampling
library(DALEX)             # for model interpretation  
library(DALEXtra)          # for extension of DALEX
theme_set(theme_minimal()) # Lisa's favorite theme
```

```{r data}
data("lending_club")
# Data dictionary (as close as I could find): https://www.kaggle.com/wordsforthewise/lending-club/discussion/170691
```


When you finish the assignment, remove the `#` from the options chunk at the top, so that messages and warnings aren't printed. If you are getting errors in your code, add `error = TRUE` so that the file knits. I would recommend not removing the `#` until you are completely finished.

## Put it on GitHub!        

From now on, GitHub should be part of your routine when doing assignments. I recommend making it part of your process anytime you are working in R, but I'll make you show it's part of your process for assignments.

**Task**: When you are finished with the assignment, post a link below to the GitHub repo for the assignment. Make sure the link goes to a spot in the repo where I can easily find this assignment. For example, if you have a website with a blog and post the assignment as a blog post, link to the post's folder in the repo. As an example, I've linked to my GitHub stacking material [here](https://github.com/llendway/ads_website/tree/master/_posts/2021-03-22-stacking).


[homework 5](https://github.com/kidakk/Advanced-Data-Science)

## Interpretable ML methods

We will once again use the lending club data that we used in the 3rd assignment. We will focus on the random forest model, which I recreate below. (Note we use this model even though the true negative rate of the training set is quite bad.)

```{r}
set.seed(494) # for reproducibility

#split data
lending_split <- initial_split(lending_club,
                               prop = .75,
                               strata = Class)

lending_training <- training(lending_split)
lending_test <- testing(lending_split)


#create recipe - including up and downsampling for model fitting
set.seed(456)
rf_recipe <- 
  recipe(Class ~ .,
         data = lending_training) %>% 
  step_upsample(Class, over_ratio = .5) %>% 
  step_downsample(Class, under_ratio = 1) %>% 
  step_mutate_at(all_numeric(), 
                 fn = ~as.numeric(.))

# create model
rf_model <- 
  rand_forest(mtry = tune(), 
              min_n = tune(), 
              trees = 100) %>% 
  set_mode("classification") %>% 
  set_engine("ranger")

# create workflow
rf_workflow <-
  workflow() %>% 
  add_recipe(rf_recipe) %>% 
  add_model(rf_model)

  grid_regular(finalize(mtry(),
                        lending_training %>%
                          select(-Class)),
               min_n(),
               levels = 3)

# create penalty grid
  rf_penalty_grid <- 
grid_regular(finalize(mtry(),
                        lending_training %>%
                          select(-Class)),
               min_n(),
               levels = 3)


# create cv samples
set.seed(494) #for reproducible 5-fold
lending_cv <- vfold_cv(lending_training,
                       v = 5)

# tune model
rf_tune <- 
  rf_workflow %>% 
  tune_grid(
    resamples = lending_cv,
    grid = rf_penalty_grid
  )

# find model with best accuracy
best_accuracy <-
  rf_tune %>% 
  select_best(metric = "accuracy")

# finalize model
rf_final <- rf_workflow %>% 
  finalize_workflow(best_accuracy) %>% 
  fit(data = lending_training)
```

1. Use functions from the `DALEX` and `DALEXtra` libraries to create a histogram and boxplot of the residuals from the training data. How do they look? Any interesting behavior?

```{r}
rf_explain <- 
  explain_tidymodels(
    model = rf_final,
    data = lending_training %>% select(-Class), 
    y = lending_training %>% 
      mutate(Class_num = as.integer(Class =="good")) %>% 
      pull(Class_num),
    label = "rf"
  )
```
```{r}
rf_mod_perf <- model_performance(rf_explain)
```

```{r}
hist_plot <- 
  plot(rf_mod_perf, 
       geom = "histogram")
box_plot <-
  plot(rf_mod_perf, 
       geom = "boxplot")

hist_plot
box_plot
```
The distribution of the residuals are right-skewed. Most of the occurrences are in the lower value cells (lower residuals appear more for the training data). The root mean square of residuals is between 0.1 and 0.2. The interquartile range is below 0.1. 

2. Use `DALEX` functions to create a variable importance plot from this model. What are the most important variables? 

```{r}
set.seed(10) #since we are sampling & permuting, we set a seed so we can replicate the results
rf_var_imp <- 
  model_parts(
    rf_explain
    )

plot(rf_var_imp, show_boxplots = TRUE)
```
The most important variable is int_rate, since the RMSE increases the most when int_rate is permuted. Other important variables are revol_until, sub_grade, annual_inc, and etc. 

3. Write a function called `cp_profile` to make a CP profile. The function will take an explainer, a new observation, and a variable name as its arguments and create a CP profile for a quantitative predictor variable. You will need to use the `predict_profile()` function inside the function you create - put the variable name there so the plotting part is easier. You'll also want to use `.data[[]]` rather than `aes()` and quote the variables. Use the `cp_profile()` function to create one CP profile of your choosing. Be sure to choose a variable that is numeric, not integer. There seem to be issues with those that I'm looking into.

For an extra challenge, write a function that will work for either a quantitative or categorical variable. 

If you need help with function writing check out the [Functions](https://r4ds.had.co.nz/functions.html) chapter of R4DS by Wickham and Grolemund.
```{r}
obs5 <- lending_club %>% 
  slice(5)

cp_profile <- function(explainer, new_obs, var){

  rf_cpp <- predict_profile(explainer = explainer, 
                          variables = var,
                          new_observation = new_obs) 
  
  rf_cpp %>% 
    filter(`_vname_` %in% c(var)) %>% 
    ggplot(aes(x = .data[[var]], #unquote this
             y = `_yhat_`)) +
    geom_line() 
  
  
}

cp_profile(rf_explain, obs5,"int_rate")
```


4. Use `DALEX` functions to create partial dependence plots (with the CP profiles in gray) for the 3-4 most important variables. If the important variables are categorical, you can instead make a CP profile for 3 observations in the dataset and discuss how you could go about constructing a partial dependence plot for a categorical variable (you don't have to code it, but you can if you want an extra challenge). If it ever gives you an error that says, "Error: Can't convert from `VARIABLE` <double> to `VARIABLE` <integer> due to loss of precision", then remove that variable from the list. I seem to have figured out why it's doing that, but I don't know how to fix it yet.

```{r}
set.seed(494) # since we take a sample of 100 obs
# This takes a while to run. 
# If we only want to examine a few variables, add variables argument to model_profile. If I remove that, it will do it for all variables and takes longer to run.
rf_pdp <- model_profile(explainer = rf_explain, 
                        variables = c("int_rate"))

plot(rf_pdp, 
     variables = c("int_rate"),
     geom = "profiles")
```

```{r}
set.seed(494) # since we take a sample of 100 obs
# This takes a while to run. 
# If we only want to examine a few variables, add variables argument to model_profile. If I remove that, it will do it for all variables and takes longer to run.
rf_pdp <- model_profile(explainer = rf_explain, 
                        variables = c("sub_grade"))

plot(rf_pdp, 
     variables = "sub_grade",
     geom = "profiles")
```

```{r}
set.seed(494) # since we take a sample of 100 obs
# This takes a while to run. 
# If we only want to examine a few variables, add variables argument to model_profile. If I remove that, it will do it for all variables and takes longer to run.
rf_pdp <- model_profile(explainer = rf_explain, 
                        variables = c("annual_inc"))

plot(rf_pdp, 
     variables = "annual_inc",
     geom = "profiles")
```


5. Choose 3 observations and do the following for each observation:  
  - Construct a break-down plot using the default ordering. Interpret the resulting graph. Which variables contribute most to each observation's prediction?  

```{r}
# obs1
new_obs1 <- lending_club %>% slice(1000) 

# Pulls together the data needed for the break-down plot
pp_rf1 <- predict_parts(explainer = rf_explain,
                          new_observation = new_obs1,
                          type = "break_down") #default

# Break-down plot
plot(pp_rf1)
```

The variable contributes the most to the observation's prediction is int_rate, which a grade of 8.49 contributes 0.05 to the average prediction. 

```{r}
# obs2
new_obs2 <- lending_club %>% slice(2000) 

# Pulls together the data needed for the break-down plot
pp_rf2 <- predict_parts(explainer = rf_explain,
                          new_observation = new_obs2,
                          type = "break_down") #default

# Break-down plot
plot(pp_rf2)
```

The variable contributes the most to the observation's prediction is addr_state, which a value of 14 contributes 0.085 to the average prediction. 

```{r}
# obs3
new_obs3 <- lending_club %>% slice(3000) 

# Pulls together the data needed for the break-down plot
pp_rf3 <- predict_parts(explainer = rf_explain,
                          new_observation = new_obs3,
                          type = "break_down") #default

# Break-down plot
plot(pp_rf3)
```
  
The variable contributes the most to the observation's prediction is int_rate, which a value of 7.39 contributes 0.057 to the average prediction. 
  
  - Construct a SHAP graph and interpret it. Does it tell a similar story to the break-down plot?  
  

```{r}
rf_shap1 <-predict_parts(explainer = rf_explain,
                        new_observation = new_obs1,
                        type = "shap",
                        B = 10 #number of reorderings - start small
)

plot(rf_shap1)
```

```{r}
rf_shap2 <-predict_parts(explainer = rf_explain,
                        new_observation = new_obs2,
                        type = "shap",
                        B = 10 #number of reorderings - start small
)

plot(rf_shap2)
```


```{r}
rf_shap3 <-predict_parts(explainer = rf_explain,
                        new_observation = new_obs3,
                        type = "shap",
                        B = 10 #number of reorderings - start small
)

plot(rf_shap3)
```

It tells a similar story.

  - Construct a LIME graph (follow my code carefully). How close is each original prediction to the prediction from the local model? Interpret the result. You can also try using fewer or more variables in the local model than I used in the example.  
  
```{r}
set.seed(2)

# NEED these two lines of code always!
# They make sure our explainer is defined correctly to use in the next step
model_type.dalex_explainer <- DALEXtra::model_type.dalex_explainer
predict_model.dalex_explainer <- DALEXtra::predict_model.dalex_explainer

lime_rf <- predict_surrogate(explainer = rf_explain,
                             new_observation = new_obs1 %>%
                               select(-Class), 
                             n_features = 5,
                             n_permutations = 1000,
                             type = "lime")

lime_rf %>% 
  select(model_r2, model_prediction, prediction) %>% 
  distinct()
```
  
  
```{r}
plot(lime_rf) +
  labs(x = "Variable")
```

The predicted value from the original random forest model (Prediction) is about 0.92.The bars show the variables ordered by weight, giving an indication of which variables are most important in the local model. In this local model, the most important variable is int_rate, following inq_last_6mths, annual_innc, total_il_high_credit_limit and total_bal_il.

6. Describe how you would use the interpretable machine learning tools we've learned (both local and global) in future machine learning projects? How does each of them help you?

I will better analyze the model output statistically, and thus more aware of how the dataset influences the outcome. I will also be able to break down the variable importance and how each variables influence the particular outcome. 

7. Save this final model using the `write_rds()` function - see the [Use the model](https://advanced-ds-in-r.netlify.app/posts/2021-03-16-ml-review/#use-the-model) section of the `tidymodels` intro for a similar example, but we're using `write_rds()` instead of `saveRDS()`. We are going to use the model in the next part. You'll want to save it in the folder where you create your shiny app. Run the code, and then add `eval=FALSE` to the code chunk options (next to the r inside the curly brackets) so it doesn't rerun this each time you knit. 

```{r, eval=FALSE}
write_rds(rf_final, 'lending_rf_final.rds')
```


## Shiny app

You are going to create an app that allows a user to explore how the predicted probability of a loan being paid back (or maybe just the predicted class - either "good" or "bad") changes depending on the values of the predictor variables.

Specifically, you will do the following:

* Set up a separate project and GitHub repo for this app. Make sure the saved model from the previous problem is also in that folder. The app needs to be created in a file called *exactly* app.R that is also in the project folder.
* At the top of the file, load any libraries you use in the app.
* Use the `read_rds()` function to load the model.  
* You may want to load some of the lending data to use to help in the design of your app. The original data are in `tidymodels` which you will also have to load in the shiny app. 
* Create a user interface (using the various `*Input()` functions) where someone could enter values for each variable that feeds into the model. You will want to think hard about which types of `*Input()` functions to use. Think about how you can best prevent mistakes (eg. entering free text could lead to many mistakes). I'd recommend using sliders and drop-downs as often as possible.
* Another part of the user interface will allow them to choose a variable (you can limit this to only the quantitative variables) where they can explore the effects of changing that variable, holding all others constant.  
* After the user has entered all the required values, the output will be a CP profile with the the predicted value for the data that was entered, indicated by a point. You may be able to use the functions from `DALEX` and `DALEXtra` or you can do some of your own coding. 
* Use the `bslib` to theme your shiny app!  
* Publish your app to [shinyapps.io](https://www.shinyapps.io/). There are instructions for doing that on my tutorial page from Intro Data Science: https://animation-and-interactivity-in-r.netlify.app/#publishing-your-app. Make sure to load ALL the libraries that you use at the top of your file. This includes the ranger library, which is used "behind the scenes" when your model is fit. If you try to publish and are unsuccessful, forgetting a library you need is the most common reason I have seen.     
* Write a paragraph or two describing your app on your website! Link to the app and your GitHub repository in your post. Include a link to your post here. 

[post](https://xuechenyu.netlify.app/posts/2021-11-20-interpretable-machine-learning-shiny-app/)

## Data Ethics: Data visualization principles

I'm hoping that the topic for this week is a review for all of you, but it's good to remind ourselves of these principles.  

**Task:**

Read both short articles in Week6. Data visualization section of [Calling Bulllshit](https://www.callingbullshit.org/syllabus.html#Visual). Were there any principles mentioned that you hadn't heard of before? What graph stood out for you as "the worst"? Did any of the graphs fool you? Or were able to overcome the bad practices (don't worry if they fool you - plently of them have fooled me, too.)? How does practicing good data visualization principles fit in with data ethics?

I have not heard of the principle of proportional link. The multiple axis on the single chart example of Thyroid cancer seems to be the worst. I don't think any of the graph fooled me. Practicing a good data visualization principle is important as data visualization is a crucial thing in data interpretation, and it will affect our understanding of the data and further implementation of data too. 

## Make this document look nicer!

Go to the top and delete the `#` from the options section and knit a final time.
